# Example R markdown document

Load tidyverse

```{r}
library(tidyverse)
```

Read data

```{r}
singleM_i <- read_delim("Data/singleM.illumina.tsv", delim = "\t") %>% 
  select(-gene, -sequence) %>% 
  mutate(sample = str_remove(sample, ".illumina.R1")) %>% 
  mutate(taxonomy = str_remove(taxonomy, "Root; ")) %>% 
  mutate(taxonomy = str_remove_all(taxonomy, "[dpcofg]__")) %>% 
  separate(taxonomy, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "; ") %>% 
  mutate_at(vars(Domain:Genus), ~replace_na(., "Unclassified")) %>% 
  filter(Domain %in% c("Archaea", "Bacteria"))

singleM_n <- read_delim("Data/singleM.nanopore.tsv", delim = "\t") %>% 
  select(-gene, -sequence) %>% 
  mutate(taxonomy = str_remove(taxonomy, "Root; ")) %>% 
  mutate(taxonomy = str_remove_all(taxonomy, "[dpcofg]__")) %>% 
  separate(taxonomy, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "; ") %>% 
  mutate_at(vars(Domain:Genus), ~replace_na(., "Unclassified")) %>% 
  filter(Domain %in% c("Archaea", "Bacteria"))
```

Aggregate at the phylum level

```{r}
singleM_i_p <- singleM_i %>% 
  group_by(across(Domain:Phylum), sample) %>% 
  summarise(coverage = sum(coverage)) %>% 
  group_by(sample) %>% 
  mutate(relabund = (coverage/sum(coverage))*100) %>% 
  ungroup()
```

Aggregate at genus level

```{r}
singleM_i_g <- singleM_i %>% 
  group_by(across(Domain:Genus), sample) %>% 
  summarise(coverage = sum(coverage)) %>% 
  group_by(sample) %>% 
  mutate(relabund = (coverage/sum(coverage))*100) %>% 
  ungroup()
```

What are the main (most abundant) taxa in our samples?  
On phylum level:

```{r}
singleM_i_p %>% 
  group_by(across(Domain:Phylum)) %>% 
  summarise(mean_relabund = mean(relabund)) %>% 
  ungroup() %>% 
  arrange(desc(mean_relabund)) %>% 
  slice(1:10)
```

On genus level:

```{r}
singleM_i_g %>% 
  group_by(across(Domain:Genus)) %>% 
  summarise(mean_relabund = mean(relabund)) %>% 
  ungroup() %>% 
  arrange(desc(mean_relabund)) %>% 
  filter(Genus != "Unclassified") %>% 
  slice(1:10)
```

What are the potential differences in community composition between the samples?  
On phylum level:

```{r}
singleM_i_p %>% 
  ggplot(aes(x = Phylum, y = relabund, fill = Domain)) +
  geom_bar(stat = "identity") +
  facet_grid(rows = vars(sample), cols = vars(Domain), space = "free", scale = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Top taxa genus level:

```{r}
TOP_GENERA <- singleM_i_g %>% 
  group_by(across(Domain:Genus)) %>% 
  summarise(mean_relabund = mean(relabund)) %>% 
  ungroup() %>% 
  arrange(desc(mean_relabund)) %>% 
  filter(Genus != "Unclassified") %>% 
  slice(1:50) %>% 
  pull(Genus)

singleM_i_g %>% 
  filter(Genus %in% TOP_GENERA) %>% 
  ggplot(aes(x = Genus, y = relabund, fill = Phylum)) +
  geom_bar(stat = "identity") +
  facet_grid(rows = vars(sample), cols = vars(Domain), space = "free", scale = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

How many unclassifed?  
On phylum level:

```{r}
singleM_i_p %>% 
  filter(Phylum == "Unclassified") %>% 
  group_by(sample) %>% 
  summarise(sum = sum(relabund))
```

On genus level:

```{r}
singleM_i_g %>% 
  filter(Genus == "Unclassified") %>% 
  group_by(sample) %>% 
  summarise(sum = sum(relabund))
```

How do the taxonomic profiles obtainted from Illumina and Nanopore data compare?

```{r}
singleM_n_p <- singleM_n %>% 
  group_by(across(Domain:Phylum), sample) %>% 
  summarise(coverage = sum(coverage)) %>% 
  group_by(sample) %>% 
  mutate(relabund = (coverage/sum(coverage))*100) %>% 
  ungroup()

bind_rows(singleM_i_p, singleM_n_p) %>% 
  ggplot(aes(x = Phylum, y = relabund, fill = Domain)) +
  geom_bar(stat = "identity") +
  facet_grid(rows = vars(sample), cols = vars(Domain), space = "free", scale = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```